---
title: "Salamander Project"
author: "Moumen Kamnaksh"
date: "11/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1

### Libraries
```{r, warning=FALSE,message=FALSE}
rm(list = ls())

# Loading all the necessary packages needed for the analyses

library(readxl)
library(lme4)
library(lmerTest)
library(merTools)
library(ggplot2)
library(tidyverse)
library(knitr)
library(ggpubr)
library(factoextra)
library(FactoMineR)
library(formatR)
```

### Read Data
```{r}

#Reading the dataset salamanderHabitat.csv which will be used 
#for the first two questions of the project

habitat_data <- read.table(file = "salamanderHabitat.csv", header = TRUE, 
                             stringsAsFactors = TRUE, sep = ",")

#Providing a summary of the dataset where minimum, the maximum, 
#the mean, the median and the amount of missing values 
#for each variable is presented

summary(habitat_data)

```

### Mixed effects model using each habitat variable separately
```{r}
#Fitting 4 separate models for each habitat variable. Year and Month are included 
#as fixed effects while Boardcode is set as random. We use function lmer() from package lme4 
#since all response variables are continuous. The summary functions provides information 
#about the restricted maximum likelihood convergence criterion, a summary of the scaled residuals, 
#the standard deviation and variance of the random effects (Boardcode and error terms) 
#and the coefficients of the fixed effects, their standard error, the degrees of freedom 
#and t-value from t-test.By loading package lmerTest 
#we also obtain the p-values of the fixed effects in the summary of the model. 
#Lastly the correlation of the fixed effects is measured. 

# Mixed effects model with SoilTemp as response variable

model1 <- lmer(SoilTemp ~ Year + Month + (1|BoardCode),
     data = habitat_data)
summary(model1)

#Mixed effects model with SoilMoisture as response variable

model2 <- lmer(SoilMoisture ~ Year + Month + (1|BoardCode),
     data = habitat_data)
summary(model2)

#Mixed effects model with CanopyCover as response variable

model3 <-  lmer(CanopyCover ~ Year + Month + (1|BoardCode),
     data = habitat_data)
summary(model3)

#Mixed effects model with LeafLitterDepth as response variable

model4 <- lmer(LeafLitterDepth ~ Year + Month + (1|BoardCode),
     data = habitat_data)
summary(model4)
```

### Visualizing the patterns between Month and the habitat variables
```{r}
#Remove missing values

habitat_data2 <- na.omit(habitat_data)

#Scatterplots showing the linear relationship between the dependent variable Month 
#and the response variables. Since the relationship between the fixed effect 
#and the response variable is linear we use geom_smooth() with method = lm. 
#We create four different ggplots where in the x axis the four months are presented 
#while on y axis we have continuous values of the habitat conditions. 

ggm1 <- ggplot(habitat_data2, aes(x = Month, y = SoilTemp)) + geom_point() + 
  geom_smooth(method = "lm")

ggm2 <- ggplot(habitat_data2, aes(x = Month, y = SoilMoisture)) + geom_point() + 
  geom_smooth(method = "lm")

ggm3 <- ggplot(habitat_data2, aes(x = Month, y = CanopyCover)) + geom_point() + 
  geom_smooth(method = "lm")

ggm4 <- ggplot(habitat_data2, aes(x = Month, y = LeafLitterDepth)) + geom_point() + 
  geom_smooth(method = "lm")

#With ggarrange() we gather all four plots in one panel

ggarrange(ggm1,ggm2,ggm3,ggm4)
```

### Visualizing the patterns between Year and the habitat variables
```{r}
#Scatterplots showing the linear relationship between the dependent variable Month 
#and the response variables. 
#We follow the same procedure as before with the only change in this case 
#in the x axis the two years 2018 and 2019 appear.

ggy1 <- ggplot(habitat_data2, aes(x = Year, y = SoilTemp)) + geom_point() + 
  geom_smooth(method = "lm")

ggy2 <- ggplot(habitat_data2, aes(x = Year, y = SoilMoisture)) + geom_point() + 
  geom_smooth(method = "lm")

ggy3 <- ggplot(habitat_data2, aes(x = Year, y = CanopyCover)) + geom_point() + 
  geom_smooth(method = "lm")

ggy4 <- ggplot(habitat_data2, aes(x = Year, y = LeafLitterDepth)) + geom_point() + 
  geom_smooth(method = "lm")

ggarrange(ggy1,ggy2,ggy3,ggy4)
```

#Part 2

### Generalized linear mixed effects model
```{r}
#We turn the variable Occupied into factor. It is a step that is required 
#since this variable is a binary one and not a numerical continuous variable. 

habitat_data$Occupied <- as.factor(habitat_data$Occupied)

#We fit a generalized linear mixed effects model since our response variable is binary. 
#We set the error distribution (family = binomial). 
#In this model SoilTemp and SoilMoisture are used as independent fixed effects while BoardCode is set as random once more. 
#The only difference with the output of an lmer() is that instead of the REML convergence criterion 
#we get information about the Akaike Information and the Bayesian Information criteria, 
#along with the log-likelihood and the deviance of the model. 
#The other difference is that instead of t-value for the fixed effects we get a z value. 
#Z value describes the position of a raw score 
#in terms of its distance from the Mean when measured in Standard deviation.

gm1 <- glmer(Occupied ~ SoilTemp + SoilMoisture + (1|BoardCode),
             data = habitat_data, family = binomial)
summary(gm1)
```

### Boxplots
```{r}

#Using SoilMoisture as a variable. Occupied is a binary variable, #therefore the two groups have to be visualized separately. 
#Boxplots can show the trends of the groups along with the variance, the mean and possible outliers.
#Therefore in x axis the two values of Occupied appear and on y axis the values of SoilMoisture are presented. 
#Two separate blue boxes appear on the final panel. 

b1 <- boxplot(SoilMoisture ~ Occupied, data = habitat_data2,
        main = "Categorized by Occupation",
        xlab = "Occupied",
        ylab = "SoilMoisture",
        col = "blue")

#Using Soiltemp as a variable. The same comments hold here as well. #The difference is that in this case on y axis, the values of the SoilTemp appear.
#Additionally, the two boxes are orange and not blue.

b2 <- boxplot(SoilTemp ~ Occupied, data = habitat_data2,
        main = "Categorized by Occupation",
        xlab = "Occupied",
        ylab = "SoilTemp",
        col = "orange")

#We gather the two boxplots in one panel

ggarrange(b1,b2)
```

### Table of aproximate range of suitable habitat conditions
```{r}
#Creating a dataset that has only the boards that are occupied

new_data <- habitat_data[which(habitat_data$Occupied == 1),]

#Creating a dataset that has only the boards that are not occupied

new_data2 <- habitat_data[which(habitat_data$Occupied == 0),]

#Removing the missing values from the dataset with no occupied boards

new_data2 <- na.omit(new_data2)

#Calculating the min, max, mean and standard deviation of Soiltemp #for boards that are occupied

st <- c(min(new_data$SoilTemp), max(new_data$SoilTemp), mean(new_data$SoilTemp),
        sd(new_data$SoilTemp))

#Calculating the min, max, mean and standard deviation of SoilMoisture 
#for boards that are occupied

sm <- c(min(new_data$SoilMoisture), max(new_data$SoilMoisture),     
        mean(new_data$SoilMoisture), sd(new_data$SoilMoisture))

#Combining these 8 values, each 4 in a separate column

t1 <- cbind(st,sm)

#Turning this vector into a data frame

t1 <- as.data.frame(t1)

#Naming the rows of the data frame 

row.names(t1) <- c("min", "max", "mean", "sd")

#Printing a table with all the information

kable(t1, digits = 10, format.args = list(scientific = -1), 
      col.names = c("SoilTemperature", "SoilMosture"),
      caption = "Range of suitable habitat conditions")

# We repeat the exact same procedure for the SoilMoisture variable 

st <- c(min(new_data2$SoilTemp), max(new_data2$SoilTemp), mean(new_data2$SoilTemp),
        sd(new_data2$SoilTemp))
sm <- c(min(new_data2$SoilMoisture), max(new_data2$SoilMoisture), 
        mean(new_data2$SoilMoisture), sd(new_data2$SoilMoisture))
t1 <- cbind(st,sm)
t1 <- as.data.frame(t1)
row.names(t1) <- c("min", "max", "mean", "sd")

kable(t1, digits = 10, format.args = list(scientific = -1), 
      col.names = c("SoilTemperature", "SoilMosture"),
      caption = "Range of unsuitable habitat conditions")

#Calculating the t-test to validate if there is a statistical significant difference 
#between the means of the two groups of Occupation for habitat variables, Soil Moisture 
#and SoilTemp, CanopyCover and LeadLitterDepth.

t.test(new_data2$SoilTemp, new_data$SoilTemp)
t.test(new_data$SoilMoisture, new_data2$SoilMoisture)
t.test(new_data2$CanopyCover, new_data$CanopyCover)
t.test(new_data2$LeafLitterDepth, new_data$LeafLitterDepth)
```

#Part 3

### Reading the pond data
```{r}
#Reading the data from the .csv format.
#Since the data are in our current working directory we only have to specify the name of the file.
#We use header = TRUE to specify that the first row contains the column names.
pond_data <- read.csv("salamanderPonds.csv", header = TRUE)

#Selecting variables of interest by using the function select().
#By using the function mutate(), we change the class of the variables PondType and Presence to factor.
pond_data <- pond_data %>%
  select(PondType, Presence, PondSurfaceArea, PondMaxDepth,
         Proximity_to_Forest, CanopyCover, 
         PredPresence, Crayfish, AquaticVeg,
         Substrate, Noon_corrected_Temp) %>%
  mutate(PondType = as.factor(PondType)) %>%
  mutate(Presence = as.factor(Presence))

#Finally, we change the levels of the Presence variable, from (0,1) to (Not present, Present)
levels(pond_data$Presence) <- c("Not present", "Present")
```

### PCA
```{r}
#Using prcomp(), we apply PCA using the recomended variables.
#By using scale. = TRUE and center = TRUE we standardize our data prior to PCA.
#This is especially usefull when we have numerical variables that are measured in different scales.
#Like PondSurfaceArea and PondMaxDepth.
#The standardized data have zero mean and standard deviation equal to one.
pond_pca <- prcomp(pond_data %>%
                     select(c(-PondType, -Presence)), scale. = TRUE, center = TRUE)

#Scree plot
#The first three components (dimensions) explain the largest percentage of the total variation
# PCA1 : ~31% VAF
# PCA2 : ~19% VAF
# PCA3 : ~17% VAF
fviz_eig(pond_pca, addlabels = TRUE)

#Extract the corresponding eigenvalues
#The first three PCs explain 66,5% of the total variation
get_eigenvalue(pond_pca)

#Individual similarity.
#Shows the individuals plotted against the first two dimensions.
#Similar individuals appear closer together.
#Individuals are coloured base on the quality of representation (cos2)
fviz_pca_ind(pond_pca,
             geom = "point",
             pointsize = 5,
             gradient.cols = c("white", "red", "blue"),
             col.ind = "cos2",
             repel = TRUE)
```

### Interpreting the first dimension
```{r}
##The following visualizations are going to help us identify any underlying patterns,
##as well as provide an interpratation for the first two principal components

#Plot the contributions of variables to the first dimension (axis = 1)
fviz_contrib(pond_pca, choice = "var", axes = 1, top = 10)

#Plot the contributions of variables to the second dimension (axis = 2)
fviz_contrib(pond_pca, choice = "var", axes = 2, top = 10)

#Biplot of individuals and variables
#Use repel = TRUE to avoid text overlap
fviz_pca_biplot(pond_pca, repel = TRUE)
```

### Investigate differences between pond types 
```{r}
##The two following biplots are going to help us identify any underlying patterns 
##with respect to Pond Type and Presence variables. 

#Create a biplot, where individuals are coloured by Pond Type
fviz_pca_biplot(pond_pca, 
                col.ind = pond_data$PondType, palette = "jco", 
                addEllipses = TRUE, label = "var",
                col.var = "black", repel = TRUE,
                legend.title = "Pond Type") 

#Create a biplot, where individuals are coloured by Presence
fviz_pca_biplot(pond_pca, 
                col.ind = pond_data$Presence, palette = "jco", 
                addEllipses = TRUE, label = "var",
                col.var = "black", repel = TRUE,
                legend.title = "Larvae Presence")
```

### Boxplots
```{r}
#Extract the first principal component 
PC1 <- pond_pca$x[, 1]

#Create boxplots
par(mfrow = c(1,2))
#Using Pond Type as a variable
boxplot(PC1 ~ pond_data$PondType,
        main = "Categorized by pond type",
        xlab = "Pond Type",
        ylab = "First proncipal component",
        col = "blue")

#Using Presence as a variable
boxplot(PC1 ~ pond_data$Presence,
        main = "Categorized by larvae presence",
        xlab = "Presence",
        ylab = "First proncipal component",
        col = "orange")
```

### Checking the assumption of normality
```{r}
#Before conducting a statistical test we will check the following assumptions:
# - Assumption of Normaliry
# - Assumption of equal variances
# If these assumptions are met, we can conduct a two samples t-test
# Otherwise a non-parametric test will be considered

#Adding PC1 as a variable in our dataset and grouping by pond type
pond_data <- pond_data %>%
mutate(PC1 = PC1) %>%
group_by(PondType)

#We have 4 groups for which we want to conduct statistical tests.
#Constructed vs Natural and Present vs Not present.
#PC1 values for Constructed ponds
PC1_Constr <- pond_data$PC1[pond_data$PondType == "Constructed"]
#PC1 values for Natural ponds
PC1_Natur <- pond_data$PC1[pond_data$PondType == "Natural"]
#PC1 values for ponds with larvae
PC1_Pres <- pond_data$PC1[pond_data$Presence == "Present"]
#PC1 values for ponds without larvae
PC1_Notpr <- pond_data$PC1[pond_data$Presence == "Not present"]

#Check assumption of normality
#A two-sample t-test assumes normal distributions in groups
#Plot the distributions
par(mfrow = c(2,2))
#Density Histogram
hist(PC1_Constr,
     breaks = 10, col = "blue", main = "Constructed Ponds",
     xlab = "PC1 loadings", freq = FALSE)
#Add a density line
lines(density(PC1_Constr),
      col = "red")

hist(PC1_Natur,
     breaks = 10, col = "blue", main = "Natural Ponds",
     xlab = "PC1 loadings", freq = FALSE)
lines(density(PC1_Natur),
      col = "red")

hist(PC1_Notpr,
     breaks = 10, col = "orange", main = "Larvae not present",
     xlab = "PC1 loadings", freq = FALSE)
lines(density(PC1_Notpr),
      col = "red")

hist(PC1_Pres,
     breaks = 10, col = "orange", main = "Larvae present",
     xlab = "PC1 loadings", freq = FALSE)
lines(density(PC1_Pres),
      col = "red")
```

### Shapiro-Wilk normality tests
```{r}
#p-values are above 0.05, therefore the null hypothesis cant be rejected
#We can assume that the four groups are normally distributed.
shapiro.test(PC1_Constr)
shapiro.test(PC1_Natur)
shapiro.test(PC1_Notpr)
shapiro.test(PC1_Pres)
```

### Checking the assumption of equal variances (F-test)
```{r}
#Test for equal variances
#p-values are above 0.05, therefore we can assume approximately equal variances
var.test(PC1 ~ PondType, data = pond_data)
var.test(PC1 ~ Presence, data = pond_data)
```

### Two sample t-tests
```{r}
#p-values < 0.001
#Reject the null hypothesis that group means are equal.
t.test(PC1_Constr, PC1_Natur, var.equal = TRUE)
t.test(PC1_Notpr, PC1_Pres, var.equal = TRUE)
```

